---
title: "CRAN Downloads"
output: github_document
always_allow_html: true
author: "Steven P. Sanderson II, MPH - Data Scientist/IT Manager"
date: "`r format(Sys.time(), '%d %B, %Y')`"
# output:
#   html_document:
#     code_folding: hide
#     highlight: tango
#     theme: flatly
#     toc: yes
#   pdf_document:
#     toc: yes
---

```{r knit_opts, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.path = "man/figures/README-",
  fig.width = 12,
  fig.height = 10
)
```

This repo contains the analysis of downloads of my `R` packages:

-   [`healthyR`](https://www.spsanderson.com/healthyR/)
-   [`healthyR.data`](https://www.spsanderson.com/healthyR.data/)
-   [`healthyR.ts`](https://www.spsanderson.com/healthyR.ts/)
-   [`healthyverse`](https://www.spsanderson.com/healthyverse/)

All of which follow the ["analyses as package"](https://rmflight.github.io/posts/2014/07/analyses_as_packages.html) philosophy this repo itself is an `R` package that can installed using `remotes::install_github()`.

I have forked this project itself from [`ggcharts-analysis`](https://github.com/thomas-neitmann/ggcharts-downloads).

While I analyze `healthyverse` packages here, the functions are written in a way that you can analyze any CRAN package with a slight modification to the `download_log` function.

This file was last updated on `r format(Sys.Date(), "%B %d, %Y")`.

```{r setup, results='hide'}
library(packagedownloads)
library(ggplot2)
library(patchwork)
library(dplyr)
library(timetk)
library(purrr)
library(knitr)
```

```{r get_data}
start_date      <- Sys.Date() - 9
end_date        <- Sys.Date() - 2
total_downloads <- download_logs(start_date, end_date)
interactive     <- FALSE
```

# Last Full Day Data
```{r last_full_day}
downloads            <- total_downloads %>% filter(date == max(date))
daily_downloads      <- compute_daily_downloads(downloads)
downloads_by_country <- compute_downloads_by_country(downloads)

p2 <- plot_cumulative_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p2 + p4 +
  plot_annotation(
    title    = "healthyverse Packages - Last Full Day",
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )

downloads %>% 
  count(package, version) %>% 
  tidyr::pivot_wider(
    id_cols       = version
    , names_from  = package
    , values_from = n
    , values_fill = 0
    ) %>% 
  kable()
```

# Current Trend

Here are the current 7 day trends for the `healthyverse` suite of packages.

```{r analysis_7_day}
downloads            <- total_downloads[date >= start_date]
daily_downloads      <- compute_daily_downloads(downloads)
downloads_by_country <- compute_downloads_by_country(downloads)

p1 <- plot_daily_downloads(daily_downloads)
p2 <- plot_cumulative_downloads(daily_downloads)
p3 <- hist_daily_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = "healthyverse Packages - 7 Day Trend",
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

# Since Inception

```{r total_data}
start_date <- as.Date("2020-11-15")

daily_downloads <- compute_daily_downloads(downloads = total_downloads)
downloads_by_country <- compute_downloads_by_country(downloads = total_downloads)

p1 <- plot_daily_downloads(daily_downloads)
p2 <- plot_cumulative_downloads(daily_downloads)
p3 <- hist_daily_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = "healthyR packages are on the Rise",
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror - Since Inception",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

# Analysis by Package

## healthyR

```{r healthyR_analysis}
start_date <- as.Date("2020-11-15")
pkg <- "healthyR"

daily_downloads <- compute_daily_downloads(
  downloads = total_downloads
  , pkg = pkg)
downloads_by_country <- compute_downloads_by_country(
  downloads = total_downloads
  , pkg = pkg)

p1 <- plot_daily_downloads(daily_downloads)
p2 <- plot_cumulative_downloads(daily_downloads)
p3 <- hist_daily_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = glue::glue("Package: {pkg}"),
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror - Since Inception",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

## healthyR.ts

```{r healthyRts_analysis}
start_date <- as.Date("2020-11-15")
pkg <- "healthyR.ts"

daily_downloads <- compute_daily_downloads(
  downloads = total_downloads
  , pkg = pkg)
downloads_by_country <- compute_downloads_by_country(
  downloads = total_downloads
  , pkg = pkg)

p1 <- plot_daily_downloads(daily_downloads)
p2 <- plot_cumulative_downloads(daily_downloads)
p3 <- hist_daily_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = glue::glue("Package: {pkg}"),
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror - Since Inception",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

## healthyR.data

```{r healthyRdata_analysis}
start_date <- as.Date("2020-11-15")
pkg <- "healthyR.data"

daily_downloads <- compute_daily_downloads(
  downloads = total_downloads
  , pkg = pkg)
downloads_by_country <- compute_downloads_by_country(
  downloads = total_downloads
  , pkg = pkg)

p1 <- plot_daily_downloads(daily_downloads)
p2 <- plot_cumulative_downloads(daily_downloads)
p3 <- hist_daily_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = glue::glue("Package: {pkg}"),
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror - Since Inception",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

## healthyverse

```{r healthyverse_analysis}
start_date <- as.Date("2020-11-15")
pkg <- "healthyverse"

daily_downloads <- compute_daily_downloads(
  downloads = total_downloads
  , pkg = pkg)
downloads_by_country <- compute_downloads_by_country(
  downloads = total_downloads
  , pkg = pkg)

p1 <- plot_daily_downloads(daily_downloads)
p2 <- plot_cumulative_downloads(daily_downloads)
p3 <- hist_daily_downloads(daily_downloads)
p4 <- plot_downloads_by_country(downloads_by_country)

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = glue::glue("Package: {pkg}"),
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror - Since Inception",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

# Some table data

```{r top_n_data_tbl}
top_n_downloads(total_downloads, 4, r_os) %>%
  set_names("OS", "Count") %>%
  kable()

top_n_downloads(total_downloads, 4, r_version) %>%
  set_names("Version","Count") %>%
  kable()

total_downloads %>% 
  count(package, version) %>% 
  tidyr::pivot_wider(
    id_cols       = version
    , names_from  = package
    , values_from = n
    , values_fill = 0
    ) %>% 
  kable()
```

# Cumulative Downloads by Package
```{r cum_pkg_dl}
p1 <- plot_cumulative_downloads_pkg(total_downloads, pkg = "healthyR")
p2 <- plot_cumulative_downloads_pkg(total_downloads, pkg = "healthyR.ts")
p3 <- plot_cumulative_downloads_pkg(total_downloads, pkg = "healthyR.data")
p4 <- plot_cumulative_downloads_pkg(total_downloads, pkg = "healthyverse")

f <- function(date) format(date, "%b %d, %Y")
patchwork_theme <- theme_classic(base_size = 24) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 14)
  )
p1 + p2 + p3 + p4 +
  plot_annotation(
    title    = "healthyR packages are on the Rise",
    subtitle = "A Summary of Downloads from the RStudio CRAN Mirror - Since Inception",
    caption  = glue::glue("Source: RStudio CRAN Logs ({f(start_date)} to {f(end_date)})"),
    theme    = patchwork_theme
  )
```

